# Results


```{r}
load(here("data", "clean", "df_process.RData"))
load(here("data", "clean", "df_rwa.RData"))
# template code for getting world data refer to  https://stackoverflow.com/questions/30706124/plotting-the-world-map-in-r
WorldData <- map_data('world') %>% filter(region != "Antarctica") %>% fortify
```


## Distribution of the number of deaths based on conflict
```{r, fig.width=10}
df_process1<- df_process %>%
              mutate(best_in_1k = best/1000, 
                     name = paste(id, ":", country))
g1 <- ggplot(df_process1, aes(x = best_in_1k)) + 
        geom_histogram(bins = 20, colour = "#80593D", fill = "steelblue") +
        scale_x_continuous(breaks= pretty_breaks()) +
        ggtitle("Distribution of deaths caused by conflict events") +
        labs(x="Best estimated deaths, in thousands (1,000) of deaths")+
        theme(plot.title = element_text(face='bold',hjust=0.5),
              plot.subtitle = element_text(hjust=0.5))
# only small values
df_process2 <- df_process %>%
              filter(best < 100)
g2<- ggplot(df_process2, aes(x = best)) + 
        geom_histogram(bins = 100, colour = "#80593D", fill = "steelblue") +
        scale_x_continuous(breaks= pretty_breaks()) +
        ggtitle("Distribution of conflicts deaths caused by conflict events",
                subtitle = "with death < 100") +
        labs(x="Best estimated deaths")+
        theme(plot.title = element_text(face='bold',hjust=0.5),
              plot.subtitle = element_text(hjust=0.5))

df_process3 <- df_process %>%
              filter(best < 20)
g3<- ggplot(df_process3, aes(x = best)) + 
        geom_histogram(bins = 20, colour = "#80593D", fill = "steelblue") +
        scale_x_continuous(breaks= pretty_breaks()) +
        ggtitle("Distribution of conflicts deaths caused by conflict events",
                subtitle = "with death < 20") +
        labs(x="Best estimated deaths")+
        theme(plot.title = element_text(face='bold',hjust=0.5),
              plot.subtitle = element_text(hjust=0.5))

df_process4<- df_process %>%
              filter(best >= 10000)%>%
              mutate(best_in_1k = best/1000, 
                     name = paste(id, ":", country))
g4<- ggplot(df_process4, aes(x = best_in_1k)) + 
        geom_histogram(bins = 20, colour = "#80593D", fill = "steelblue", boundary=10000) +
        scale_x_continuous(breaks= pretty_breaks()) +
        ggtitle("Distribution of conflicts deaths caused by conflict events",
                subtitle = "with death >= 10,000") +
        labs(x="Best estimated deaths, in thousands (1,000) of deaths")+
        theme(plot.title = element_text(face='bold',hjust=0.5),
              plot.subtitle = element_text(hjust=0.5))

ggarrange(g1, g2, g3, g4, ncol = 2, nrow = 2)
```

As shown in the top-left plot, the dataset is highly skewed and imbalanced.  When we focus on the data with death less than 100 people, we can see the data are more concentrated in <20 regions. Most conflict events have less than five deaths.


```{r, fig.width=8}
options(scipen = 999)
df_process5<- df_process %>%
              filter(best >= 5000)%>%
              mutate(best_in_1k = best/1000, 
                     name = paste(id, ":", country))

ggplot(df_process5, aes(x = best_in_1k, y = reorder(name, best))) +
      geom_point(color = "steelblue") +
      labs(x="Best estimated deaths, in thousands (1,000) of deaths",
           y="Confict Event ID: Country") + 
      ggtitle("Confict events causing most deaths")+
        theme(plot.title = element_text(face='bold',hjust=0.5))
```

The above Cleveland plot shows top conflict events which caused most deaths ($\geq$ 5,000), there is a conflict event happened in Rwanda caused approximately 300,000 deaths, which is much more than other conflicts event. Also, most of those events causing significant fatalities were taken place in Rwanda. 

```{r}
# The conflict with largest death
df_2<- head(df_process[order(-df_process$best),], 10)[c("id", "side_a", "side_b", "year", "best")]
knitr::kable(df_2)
```

After an investigation of those conflict events, we found out most of them are between the Government of Rewanda and civilians, which was related to the [Rwanda genocide](https://en.wikipedia.org/wiki/Rwandan_genocide){target="_blank"} in 1994.


## Relation between conflict and time

```{r, fig.width=5, fig.height=3}
df_count <- df_process %>%
            group_by(year) %>%
            tally()
ggplot(df_count, aes(x = year, y = n)) +
      geom_bar(stat = "identity", fill = "steelblue") +
      ggtitle("Number of conflict events based on year") +
      labs(y = "number of conflict events") +
      scale_x_continuous(breaks = round(seq(1990,2018, by = 5),1))+
        theme(plot.title = element_text(face='bold',hjust=0.5))
```

As shown in the bar chart, the number of conflict events that occur in each year tends to increase from 1998 to 2018. The number of conflicts is lowest in 1997.

```{r}
# get conflicts' deaths by year
df_sum <- df_process %>%
            group_by(year) %>%
            summarise_at(vars(best), funs(sum)) %>%
              mutate(best_in_1k = best/1000)

# get conflicts' deaths by year without the Rwanda genocide
df_sum2 <- df_process %>%
            filter(!(df_process$id %in% df_rwa$id)) %>%
            group_by(year) %>%
            summarise_at(vars(best), funs(sum)) %>%
              mutate(best_in_1k = best/1000)
```
```{r, fig.width=10, fig.height=3}
g5<- ggplot(df_sum, aes(x = year, y = best_in_1k)) +
        geom_bar(stat = "identity", fill = "steelblue") +
        ggtitle("Deaths caused by conflicts based on year") +
        labs(y = "number of deaths, in thousands (1,000)") +
        scale_x_continuous(breaks = round(seq(1990,2018, by = 5),1))+
        theme(plot.title = element_text(face='bold',hjust=0.5))

g6<- ggplot(df_sum2, aes(x = year, y = best_in_1k)) +
        geom_bar(stat = "identity", fill = "steelblue") +
        ggtitle("Deaths caused by conflicts based on year",
                subtitle="without the data of the Rwanda genocide") +
        labs(y = "number of deaths, in thousands (1,000)") + 
        scale_x_continuous(breaks = round(seq(1990,2018, by = 5),1))+
        theme(plot.title = element_text(face='bold',hjust=0.5),
              plot.subtitle = element_text(hjust=0.5))

ggarrange(g5, g6, ncol = 2, nrow = 1)
```

The conflict-caused deaths are extremely high in 1994, as shown in the left plot. As mentioned in the previous point,  a high number of deaths were probably related to the Rwanda genocide. When we filter out the data related to the Rwanda genocide, we get the plot on the right. The conflict-caused death per year decreased from 1989 until 2005, after 2005, the number tends to increase.

```{r, fig.width=8, fig.height=3}
df_type <- df_process %>%
              group_by(year, type_of_violence) %>%
              tally()

ggplot(df_type, aes(fill=type_of_violence, x=year, y=n)) + 
    geom_bar(position="dodge", stat="identity") + 
    theme(legend.position='bottom') +
    labs(x="year", y="number of conflict events", fill="violence type") + 
    ggtitle("Number of conflict events based on year and type") +
    scale_fill_viridis_d()+
    scale_x_continuous(breaks = round(seq(1990,2018, by = 5),1)) +
        theme(plot.title = element_text(face='bold',hjust=0.5))
```

Most conflict events are state-based conflict, i.e., the conflict involved with "[a] government of a state or any opposition organization or alliance of organizations" [[source](https://ucdp.uu.se/downloads/ged/ged191.pdf){target="_blank"}, pp.28-29], which is higher in number comparing to the other two types of conflict. Before 2018, there are more one-side conflicts (against civilians) than non-state conflicts (between two armed groups).  But in 2018, there are more non-state conflicts compare to one-side conflict. 

If you are interested in the detailed and formal definition of these three types of conflict, please refer to this [source](https://ucdp.uu.se/downloads/ged/ged191.pdf){target="_blank"}, pp. 28-31.

We want to check further how many civilians deaths each year, compare to the deaths of non-civilians caused by conflicts.

```{r, fig.width=10, fig.height=3.8}
library(tidyverse)

df_civ <- df_process %>%
              group_by(year) %>%
              summarise_at(vars(best, deaths_civilians), funs(sum))%>% 
              mutate(best = best/1000, deaths_civilians = deaths_civilians/1000)%>%
              mutate(deaths_non_civilians = best-deaths_civilians)%>% 
              select(year, deaths_civilians, deaths_non_civilians) %>% 
              gather(`deaths_civilians`, `deaths_non_civilians`, 
                     key = "section", value = "deaths") 

df_civ2 <- df_process %>%
            filter(!(df_process$id %in% df_rwa$id)) %>%
            group_by(year) %>%
            summarise_at(vars(best, deaths_civilians), funs(sum))%>% 
            mutate(best = best/1000, deaths_civilians = deaths_civilians/1000)%>%
            mutate(deaths_non_civilians = best-deaths_civilians)%>% 
            select(year, deaths_civilians, deaths_non_civilians) %>% 
            gather(`deaths_civilians`, `deaths_non_civilians`, 
                   key = "section", value = "deaths")

              
gc1 <- ggplot(df_civ, aes(fill=section, y=deaths, x=year)) +
        geom_bar(position="stack", stat="identity") + 
        theme(legend.position='bottom') +
        labs(x="year", y="number of deaths, in thousands (1,000)", fill="civilian death?") + 
        ggtitle("Deaths caused by conflicts based on year") +
        scale_x_continuous(breaks = round(seq(1990,2018, by = 5),1)) +
        scale_fill_viridis_d()+
        guides(fill = guide_legend(reverse=TRUE)) +
        theme(plot.title = element_text(face='bold',hjust=0.5))

gc2 <- ggplot(df_civ2, aes(fill=section, y=deaths, x=year)) +
        geom_bar(position="stack", stat="identity") + 
        theme(legend.position='bottom') +
        labs(x="year", y="number of deaths, in thousands (1,000)", fill="civilian death?") + 
        ggtitle("Deaths caused by conflicts based on year",
                subtitle="without the data of the Rwanda genocide") +
        scale_x_continuous(breaks = round(seq(1990,2018, by = 5),1)) +
        scale_fill_viridis_d()+ 
        guides(fill = guide_legend(reverse=TRUE)) +
        theme(plot.title = element_text(face='bold',hjust=0.5),
              plot.subtitle = element_text(hjust=0.5))

ggarrange(gc1, gc2, ncol = 2, nrow = 1)
```

```{r, fig.width=10, fig.height=6}
df_civ3 <- df_process %>%
              group_by(year) %>%
              summarise_at(vars(best, deaths_civilians), funs(sum))%>% 
              mutate(deaths_civilians = 100*deaths_civilians/best)%>% 
              mutate(deaths_non_civilians = 100-deaths_civilians)%>% 
              select(year, deaths_civilians, deaths_non_civilians) %>% 
              gather(`deaths_civilians`, `deaths_non_civilians`, 
                     key = "section", value = "deaths")
df_civ4 <- df_process %>%
            filter(!(df_process$id %in% df_rwa$id)) %>%
            group_by(year) %>%
            summarise_at(vars(best, deaths_civilians), funs(sum))%>% 
            mutate(deaths_civilians = 100*deaths_civilians/best)%>% 
            mutate(deaths_non_civilians = 100-deaths_civilians)%>% 
            select(year, deaths_civilians, deaths_non_civilians) %>% 
            gather(`deaths_civilians`, `deaths_non_civilians`, 
                   key = "section", value = "deaths")

gc3 <- ggplot(df_civ3, aes(fill=section, y=deaths, x=year)) +
        geom_bar(position="stack", stat="identity") + 
        theme(legend.position='bottom') +
        labs(x="year", y="propotion of total deaths", fill="civilian death?") + 
        ggtitle("Deaths caused by conflicts based on year") +
        coord_flip() +
        scale_x_reverse(breaks = round(seq(1990,2018, by = 5),1)) +
        scale_fill_viridis_d()+
        guides(fill = guide_legend(reverse=TRUE)) +
        theme(plot.title = element_text(face='bold',hjust=0.5))

gc4 <- ggplot(df_civ4, aes(fill=section, y=deaths, x=year)) +
        geom_bar(position="stack", stat="identity") + 
        theme(legend.position='bottom') +
        labs(x="year", y="propotion of total deaths", fill="civilian death?") + 
        ggtitle("Deaths caused by conflicts based on year",
                subtitle="without the data of the Rwanda genocide") +
        coord_flip() +
        scale_x_reverse(breaks = round(seq(1990,2018, by = 5),1)) +
        scale_fill_viridis_d()+
        guides(fill = guide_legend(reverse=TRUE))  +
        theme(plot.title = element_text(face='bold',hjust=0.5),
              plot.subtitle = element_text(hjust=0.5))

ggarrange( gc3, gc4, ncol = 2, nrow = 1)
```

```{r}
ttt <- df_process %>%
        filter(year ==1996)

df_3 <- head(ttt[order(-ttt$deaths_civilians),], 10)[c("id", "side_a", "side_b", "year", "deaths_civilians", "country")]
knitr::kable(df_3)
```

From the above plots, the top-left one, we can see that there are large civilians death in conflicts happened in 1994, which potentially because of the Rwanda genocide; after removing the records regarding the Rwanda genocide, the civilian deaths in 1994 was similar with other years. 

The bottom two plots show the proportion of civilian deaths caused by conflicts, by comparing the left and right plots, we can see that the Rwanda genocide is the reason of leading extreme high proportion of civilian deaths.  In most of the years, the percentage of civilian deaths is around 25%, except in the year 1994 and the year 1996. After further checking, the reason for the high proportion of civilian deaths in 1996 is probably due to the [First Congo War](https://en.wikipedia.org/wiki/First_Congo_War){target="_blank"}.

## Patterns between geographical location and type of violence

Fristly, we want to briefly see the pattern between region and type of violence. To explore the pattern between region and type of violence, which are both categoical variables, the best way is to create a mosaic plot to see if there exists any patterns betweem them. We create a mosaic plot to show the patterns of types of violence under each different region, which means that, in the plot, region is splited first and then type of violence is splited.

```{r,fig.width=8}
p2 <- df_process %>% select(country,region,type_of_violence)

m <- ggplot(p2)+
  geom_mosaic(aes(x=product(type_of_violence,region),fill=type_of_violence,offset=0.02))+
  ggtitle("Mosaic plot for type of violence in each region")+
  xlab("Regions")+
  ylab("Type of violence")+
  scale_fill_viridis_d()+
  theme_classic() +
  theme(axis.line = element_blank(),axis.ticks = element_blank(),legend.position = "none")+
  theme(plot.title = element_text(face='bold',hjust=0.5))
m
```

From the mosaic plot above, the percentages of one-side violence and non-state conflict are both lower than percentage of state-based conflict. State-based conflict are the most frequent type of violence among all regions, as state-based conflict is the most conventional and common among all types of violence, which is conflict between two legitimate players, namely state actors or the governments of the states. State-based conflict is a very common type of violence around the world. 

Moreover, compared with Asia, Europe and Middle East, Africa and American have more percentage of non-state conflict. For Africa, there are a lot of private military companies and especially, terrorist organizations, such as Islamic State (also known as ISIS) in Libya, Boko Haram in Nigeria and al Shabab in Somalia. These military companies and terrorist organizations are senseless and they used extremely brutal method, such as suicide bombing, to hurt innocent civilians in Africa. For Americas, the reason why the percentage of non-state conflict is so high is that, in South America and Mexico, there are so many drug cartels which is the criminal organization with intention of supplying drug trafficking operations. There are always conflicts between different drug cartels.

In addition, in the graph, the total area of Africa and total area of Asia are much larger than those of Americas, Europe and Middle East, which shows that the total numbers of conflicts in Africa and Asia are greater. For Africa, as we discussed before, the terrorist organizations and private military companies are the fundational cause of such large number of conflict. However, in Asia, the reason why the number of conflict is so high remains unclear. We need to create a separate bar chart for country in Asia to see what happened.

```{r,fig.width=10}
as <- p2 %>% filter(region=="Asia")
as_new <- as %>% group_by(country,type_of_violence) %>% summarise(Freq=n())
a_3 <- ggplot(as_new, aes(fill=type_of_violence, y=Freq, x=country)) + 
    geom_bar(position="stack", stat="identity")+
    scale_fill_viridis_d()+
    ggtitle("Stack bar chart for regions in Asia, with different type of violence(1989-2018)")+
    xlab("Country in Asia")+
    ylab("Frequency of conflict (thousands of views)")+
    labs(fill="Type of violence")+
    scale_y_continuous(labels=c("0","10","20","30"))+
    theme(plot.title = element_text(face='bold',hjust=0.5)
          ,legend.position = "bottom")+
    coord_flip()
a_3

```

As we can see from the graph above, in Asia, Afghanistan has a extrmemely large number of conflicts happened, and most of these conflicts are state-based conflict. War in Afghanistan is well-known around the world. US government engaged into War in Afghanistan to wipe out Taliban government because the Taliban government refused to hand over terrorist leader Osama bin Laden in the wake of al-Qaeda’s September 11, 2001. In Afghanistan, most conflicts are state-based conflicts, which are conflicts between US and Taliban government. The war had last for eighteen years and obviously, numerous of conflicts happened during 18 years. Now we know that war in Afghanistan from 2001 till now causes relatively high total number of conflicts in Asia.

After anaysis of relation between region and type of violence, we want to explore more deeply to countries in different representative countries , just as we analysis the Afghanistan above. We decide to use the best way to explore spatial patterns which is to create a Choropleth plot to see distribution of different type of violence for each country among the world.

```{r,fig.width=15,fig.height=7}
temp6 <- df_process %>% filter(best>0) %>% arrange((best))
ggplot() + 
    geom_map(data = WorldData, map = WorldData,
                  aes(map_id=region),
                  fill = "white", colour = "gray", size=0.5) + 
    geom_point(data=temp6, aes(x=longitude, y=latitude, color=type_of_violence),
               size=0.8, alpha=0.5) +
    coord_map("rectangular", lat0=0, xlim=c(-180,180), ylim=c(-60, 90)) +
    labs(color="Type of violence", title="Choropleth map for different type of violence in countries around the world(1989-2018)",
         x="", y="") +
    scale_color_viridis_d()+
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),axis.text.x = element_blank(),axis.text.y = element_blank(),axis.ticks.x = element_blank(),axis.ticks.y = element_blank(),plot.title = element_text(face='bold',hjust=0.5,size=20),legend.position = "bottom",legend.text = element_text(size=12))

```
From the Choropleth map above, we can re-verify our finding before, it is easy to see that the number of yellow color dots which represent the state-based conflict, much more than the number of other two color dots. In other words, the conventional state-based conflict is widely happened around the wolrd compared to other two types of violence. 

Besides what we find before, we can still see a lot of patterns between countries and type of violence. Because of numerous patterns and large number of conflicts, we will only choose some of them to discuss.

Especially, there are a lot of state-based conflicts in Philippines, as we can see from the graph. This is known as Communist rebellion in the Philippines. It is an ongoing conflict between the Government of the Philippines and the communist coalition of the Communist Party of the Philippines(CPP) from 1969 to present. The long-lasting conflict betweem them can be divided into numerous small conflict which are shown in the map. We can use statistical method to see if our finding is true.
```{r,fig.width=8}
ph <- df_process %>% filter(country=="Philippines"&type_of_violence=="state-based conflict")
ph_2 <- ph %>% group_by(side_b) %>%summarise(no_rows = length(side_b))
ph_2 <- ph_2 %>%
          mutate(side_b=replace(side_b, side_b=="Military faction (forces of Honasan, Abenina & Zumel)", "Military faction\n (Honasan, Abenina & Zumel)"))


ggplot(ph_2,aes(x=side_b,y=no_rows))+
  geom_bar(stat = "identity", fill = "steelblue")+
  labs(x="Side b",y="Frequency of state-based conflict",title="State-based conflict happened in Philippines")+
  theme_grey(12)+
  coord_flip()+
    theme(plot.title = element_text(face='bold',hjust=0.5)
          ,legend.position = "bottom")
```
From the bar chart above, we can confirm that our finding is true that most of state-based conflicts on the map is the Comummist Rebellion started by CPP in the Philippines.

Meanwhile, most of conflict Mexico are non-state conflict. Our former finding is that the cause of such a great number of non-state conflicts is about the drug cartel. Therefore, we filter the data and see if it is caused by drug cartels.
```{r}
mx <- df_process %>% filter(country=="Mexico"&type_of_violence=="non-state conflict")
a <- list(unique(mx['side_a']))
b <- list(unique(mx['side_b']))
print("List of name of side A:")
print(a)
print("List of name of side B:")
print(b)

```
From the lists of two sides in non-state conflicts in Mexico, we can see that the non-state conflicts are mostly caused by fights between drug cartels.

In addition, for Peru in south america, there are a lot of yellow dots which represent state-based conflicts. It represents the internal war in Peru between the Government of Peru and the Communist Party of Peru-Shining Path/Túpac Amaru Revolutionary Movement(MRTA). The internal war of Peru last for 20 years. Then, we also filter the orginal dataset to see if our finding is true.
```{r, fig.height=3}
pe <- df_process %>% filter(country=="Peru"&type_of_violence=="state-based conflict")
pe_2 <- pe %>% group_by(side_b) %>%summarise(no_rows = length(side_b))
ggplot(pe_2,aes(x=side_b,y=no_rows))+
  geom_bar(stat = "identity", fill = "steelblue")+
  labs(x="Side b",y="Frequency of state-based conflict",title="State-based conflict happened in Peru")+
  theme_grey(12)+
  coord_flip()+
    theme(plot.title = element_text(face='bold',hjust=0.5)
          ,legend.position = "bottom")
```
From the graph above, we can see that side B with highest frequency is Sendero Luminoso, which is the spanish of Shining Path. Judging from this, we can confirm that the high number of state-based conflicts in Peru is caused by internel war in Peru between Government of Peru and Communist Part of Peru-Shining Path.

Overall, all around the world, most of regions have a dominating number of state-based conflict, except for some representative regions, such as Mexican and Central Africa. Theses exceptions are caused by special backgrounds, such as prevalence of drug cartels or rise of terrorist orgainzations.
